A Fast Unified Model for Parsing and Sentence Understanding

### 2016-05-27 writing/hard_stack_paper/paper.tex
Fix transition notation; move tuning out of appendix.

\todo{[SB] Use $a$ for transitions.}

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf t = (t_0, \dots, t_{T-1})$, where each $t_t \in \{\shift, \reduce\}$ specifies one step of the parsing process. In general a parser may also generate these transitions on the fly as it reads the tokens. It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $2N - 1$ transitions.

The parser uses two auxiliary data structures: a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. The parser is initialized with the stack empty and the buffer containing the tokens $\mathbf x$ of the sentence in order. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. It next proceeds through the transition sequence, where each transition $t_t$ selects one of the two following operations. Below, the $\mid$ symbol denotes the \textit{cons} (concatenation) operator. We arbitrarily choose to always \textit{cons} on the left in the notation below.
\vspace*{-1ex} % dirty spacing trick that somehow seemed needed right now
%TRIMMABLE
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it on to the top of the stack.
\vspace*{-1ex}
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back on to the stack.
\end{description}

@@

We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}. For each input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element. To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$. We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack. To retrieve the top two elements of the stack, we look up the two rows indicated by the final two elements of the queue $Q$. These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack. Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation. It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.

@@

\begin{algorithm}[t]
\caption{The thin stack algorithm}
\label{alg:thin-stack}
\begin{algorithmic}[1]
  \Function{Step}{bufferTop, op, $t$, $S$, $Q$}
    \If{op = \shift}
      \State $S$[$t$] := bufferTop
    \ElsIf{op = \reduce}
      \State right := $S$[$Q$.pop()]
      \State left := $S$[$Q$.pop()]
      \State $S$[$t$] := \Call{Compose}{left, right}
    \EndIf
    \State $Q$.push($t$)
  \EndFunction
\end{algorithmic}
\end{algorithm}

\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf t$  and a token sequence $\mathbf x$ as its inputs for each sentence. The token sequence is simply the words in the sentence in order. $\mathbf t$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift\ transition and each `)' as a \reduce\ transition, as here:

\vspace{0.5em}
{\noindent\small
{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
{$\mathbf t$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
{$\mathbf x$}: the, cat, sat, down
}

\paragraph{Handling variable sentence lengths} For any sentence model to be trained with batched computation, it is necessary to pad or crop sentences to a fixed length. We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI\@. Transition sequences $\mathbf t$ are cropped at the left or padded at the left with \shift s. Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf t$, and can then be padded with empty tokens at the right to meet the desired length $N$.

@@

@@

\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}_{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_t^p, \mathcal{L}_t^h\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters. The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
\begin{equation}
\begin{split}
\mathcal{L}_{\text{m}} = &\mathcal{L}_{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_{\text{t}}^{p} + \mathcal{L}_{\text{t}}^{h}) + \lambda \norm{\theta}^2_2
\end{split}
\end{equation}

@@

A supplement discusses hyperparameter tuning.
% TODO for arXiv: Make to appendix.

>>>>

@@

A shift-reduce parser accepts a sequence of input tokens $\mathbf x = (x_0, \dots, x_{N-1})$ and consumes transitions $\mathbf a = (a_0, \dots, a_{T-1})$, where each $a_t \in \{\shift, \reduce\}$ specifies one step of the parsing process. In general a parser may also generate these transitions on the fly as it reads the tokens. It proceeds left-to-right through a transition sequence, combining the input tokens $\mathbf x$ incrementally into a tree structure. For any binary-branching tree structure over $N$ words, this requires $T=2N-1$ transitions and $2N$ states.

The parser uses two auxiliary data structures: a stack $S$ of partially completed subtrees and a buffer $B$ of tokens yet to be parsed. The parser is initialized with the stack empty and the buffer containing the tokens $\mathbf x$ of the sentence in order. Let $\langle S, B \rangle = \langle \emptyset, \mathbf x \rangle$ denote this starting state. It next proceeds through the transition sequence, where each transition $a_t$ selects one of the two following operations. Below, the $\mid$ symbol denotes the \textit{cons} (concatenation) operator. We arbitrarily choose to always \textit{cons} on the left in the notation below.
\vspace*{-1ex} % dirty spacing trick that somehow seemed needed right now
%TRIMMABLE
\begin{description}
  \item[\shift:] $\langle S, x \mid B \rangle \to \langle x \mid S, B \rangle$. This operation pops an element from the buffer and pushes it on to the top of the stack.
\vspace*{-1ex}
  \item[\reduce:] $\langle x \mid y \mid S, B \rangle \to \langle (x, y) \mid S, B \rangle$. This operation pops the top two elements from the stack, merges them into a binary tree with children $(x, y)$, and pushes the result back on to the stack.
\end{description}

@@

We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}. For each input sentence, we represent the stack with a single $(T + 1) \times D$ matrix $S$. Each row $S_t$ represents the top of the actual stack at timestep $t$. At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element. To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$. We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack. To retrieve the top two elements of the stack, we look up the two rows indicated by the final two elements of the queue $Q$. These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack. Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation. It operates on the single $(T + 1) \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.

@@

\begin{algorithm}[t]
\caption{The thin stack algorithm}
\label{alg:thin-stack}
\begin{algorithmic}[1]
  \Function{Step}{bufferTop, $a$, $t$, $S$, $Q$}
    \If{$a$ = \shift}
      \State $S$[$t$] := bufferTop
    \ElsIf{$a$ = \reduce}
      \State right := $S$[$Q$.pop()]
      \State left := $S$[$Q$.pop()]
      \State $S$[$t$] := \Call{Compose}{left, right}
    \EndIf
    \State $Q$.push($t$)
  \EndFunction
\end{algorithmic}
\end{algorithm}

\paragraph{Preparing the data} At training time, SPINN requires both a transition sequence $\mathbf a$  and a token sequence $\mathbf x$ as its inputs for each sentence. The token sequence is simply the words in the sentence in order. $\mathbf a$ can be obtained from any constituency parse for the sentence by first converting that parse into an unlabeled binary parse, then linearizing it (with the usual in-order traversal), then taking each word token as a \shift\ transition and each `)' as a \reduce\ transition, as here:

\vspace{0.5em}
{\noindent\small
{\bf Unlabeled binary parse:} ( ( the cat ) ( sat down ) )\\
{$\mathbf a$}: \shift, \shift, \reduce, \shift, \shift, \reduce, \reduce\\
{$\mathbf x$}: the, cat, sat, down
}

\paragraph{Handling variable sentence lengths} For any sentence model to be trained with batched computation, it is necessary to pad or crop sentences to a fixed length. We fix this length at $N = 25$ words, longer than about 98\% of sentences in SNLI\@. Transition sequences $\mathbf a$ are cropped at the left or padded at the left with \shift s. Token sequences $\mathbf x$ are then cropped or padded with empty tokens at the left to match the number of \shift s added or removed from $\mathbf a$, and can then be padded with empty tokens at the right to meet the desired length $N$.

@@

\begin{table*}[t]\small
\begin{center}
\begin{tabular}{lrlrrrr}
\toprule
Param.     & Range & Strategy        & RNN       & SPINN-PI-NT   & SPINN-PI  & SPINN \\
\midrule 
Initial LR & 2e-4--2e-2 & \textsc{log} & 5e-3  & 3e-4 & 7e-3  & 2e-3\\
L2 regularization $\lambda$ & 8e-7--3e-5   & \textsc{log} & 4e-6  & 3e-6   & 2e-5  & 3e-5\\
Transition cost $\alpha$  & 0.5--4.0 & \textsc{lin} & --- & --- & ---  & 3.9    \\
Embedding transformation dropout keep rate & 80--95\% & \textsc{lin} & --- & 83\% & 92\%  & 86\%\\
Classifier MLP dropout keep rate & 80--95\% & \textsc{lin} & 94\%  & 94\%   & 93\%  & 94\%\\
Tracking LSTM size $D_\text{tracking}$ & 24--128 & \textsc{log} & --- & --- & 61  & 79\\
Classifier MLP layers & 1--3 & \textsc{lin} & 2 & 2 & 2 & 1\\
\bottomrule
\end{tabular}
\end{center}
\caption{
\label{tab:hyperparams}
Hyperparameter ranges and values. \ii{Range} shows the hyperparameter ranges explored during random search. \ii{Strategy} indicates whether sampling from the range was uniform, or log--uniform.
}
\end{table*}

@@

\paragraph{The objective function} Our objective combines a cross-entropy objective $\mathcal{L}^{\text{s}}$ for the SNLI classification task, a cross-entropy objective $\{\mathcal{L}_t^{\text{p}}, \mathcal{L}_t^{\text{h}}\}$ for each parsing decision for each of the sentences at each step $t$, and an L2 regularization term on the trained parameters. The terms are weighted using the tuned hyperparameters $\alpha$ and $\lambda$:
\begin{equation}
\begin{split}
\mathcal{L}_{\text{m}} = &\mathcal{L}^{\text{s}} + \alpha \sum_{t=0}^{T-1} (\mathcal{L}_t^{\text{p}} + \mathcal{L}_t^{\text{h}}) + \lambda \norm{\theta}^2_2
\end{split}
\end{equation}

@@

We use random search to tune the hyperparameters of each model, setting the ranges for search for each hyperparameter heuristically (and validating the reasonableness of the ranges on the development set), and then launching eight copies of each experiment each with newly sampled hyperparameters from those ranges. Table~\ref{tab:hyperparams} shows the hyperparameters used in the best run of each model.
